### 2025-08-01 data-visualization.qmd
20250801 - shapes and tweaks

The `ggplot2` package [@R-ggplot2] is based on the grammar of graphics [@ggplot3rdEd], which is the principle that you can build any graph from three components: 1) the data, 2) the coordinate system and aesthetics, and 3) the geometries ("geoms")—visual marks that represent data points.
The general syntax for plots created using the `ggplot2` package [@R-ggplot2] is:

@@

### Geometries {#sec-ggplot2SyntaxGeoms}

Here is a list of common geometrics ("geoms") available in `ggplot2` (there are many others):

@@

Common point shapes include circles, triangles, squares, etc.
A full list of point shapes available is provided here: <https://ggplot2.tidyverse.org/articles/ggplot2-specs.html> (archived at <https://perma.cc/N9ED-6LUS>).
To specify the point shape, you can use the `shape` parameter, as in @fig-pointShape.

>>>>

The `ggplot2` package [@R-ggplot2] is based on the grammar of graphics [@ggplot3rdEd], which is the principle that you can build any graph from three components: 1) the data, 2) the coordinate system (and aesthetic attributes), and 3) the geometric objects ("geoms")—visual marks that represent data points (e.g., points, lines, and bars).
The general syntax for plots created using the `ggplot2` package [@R-ggplot2] is:

@@

You can layer multiple geoms in the same plot, layer by layer.
You can have a different data set and aesthetics associated with each geom, by specifying the `data` and `mapping` arguments for the function of that particular geom.
For an example of a plot that uses different data sets for different geoms, see @fig-bootstrapRegressionCoefficients (@sec-multipleRegressionRobustRegression).

### Geometric Objects {#sec-ggplot2SyntaxGeoms}

Here is a list of common geometric objects ("geoms") available in `ggplot2` (there are many others):

@@

Common point shapes include circles, triangles, squares, etc.
A full list of point shapes available is provided here: <https://ggplot2.tidyverse.org/articles/ggplot2-specs.html> (archived at <https://perma.cc/N9ED-6LUS>) and is in @fig-pointShapeOptions.

::: {#fig-pointShapeOptions}
![](images/pointShapes.png){width=100% fig-alt="Available Point Shapes in ggplot2."}

Available Point Shapes in ggplot2.
:::

To specify the point shape, you can use the `shape` parameter, as in @fig-pointShape.

### 2025-08-01 multiple-regression.qmd
20250801 - shapes and tweaks

ggplot(
  quadraticTermsRegressionModel_tidy,
  aes(
    x = term,
    y = estimate)) +
  geom_point() +
  geom_errorbar(
    aes(
      ymin = conf.low,
      ymax = conf.high),
    width = 0.2) +
  coord_flip() + # flip axes for readability
  labs(
    title = "Regression Coefficients with 95% CI",
    x = "Predictor",
    y = "Coefficient Estimate") +
  theme_minimal()
```

@@

ggplot(
  quadraticTermsRegressionModelStandardized_tidy,
  aes(
    x = term,
    y = estimate)) +
  geom_point() +
  geom_errorbar(
    aes(
      ymin = conf.low,
      ymax = conf.high),
    width = 0.2) +
  coord_flip() + # flip axes for readability
  labs(
    title = "Standardized Regression Coefficients with 95% CI",
    x = "Predictor",
    y = "Standardized Coefficient Estimate") +
  theme_minimal()
```

@@

ggplot(
  boot_coefs,
  aes(estimate)) +
  geom_histogram(
    fill = "gray80",
    color = "black") +
  facet_wrap(
    ~ term,
    scales = "free") +
  geom_vline(
    aes(
      xintercept = .lower),
    data = percentile_intervals,
    col = "blue",
    linetype = "dashed") +
  geom_vline(
    aes(
      xintercept = .upper),
    data = percentile_intervals,
    col = "blue",
    linetype = "dashed") +
  labs(
    title = "Bootstrap Distributions of Coefficients",
    x = "Coefficient Estimate",
    y = "Count") +
  theme_classic()
```

>>>>

ggplot2::ggplot(
  data = quadraticTermsRegressionModel_tidy,
  aes(
    x = term,
    y = estimate)) +
  geom_point() +
  geom_errorbar(
    aes(
      ymin = conf.low,
      ymax = conf.high),
    width = 0.2) +
  coord_flip() + # flip axes for readability
  labs(
    title = "Regression Coefficients with 95% CI",
    x = "Predictor",
    y = "Coefficient Estimate") +
  theme_minimal()
```

@@

ggplot2::ggplot(
  data = quadraticTermsRegressionModelStandardized_tidy,
  aes(
    x = term,
    y = estimate)) +
  geom_point() +
  geom_errorbar(
    aes(
      ymin = conf.low,
      ymax = conf.high),
    width = 0.2) +
  coord_flip() + # flip axes for readability
  labs(
    title = "Standardized Regression Coefficients with 95% CI",
    x = "Predictor",
    y = "Standardized Coefficient Estimate") +
  theme_minimal()
```

@@

ggplot(
  data = boot_coefs,
  aes(x = estimate)) +
  geom_histogram(
    fill = "gray80",
    color = "black") +
  facet_wrap(
    ~ term,
    scales = "free") +
  geom_vline(
    data = percentile_intervals,
    aes(
      xintercept = .lower),
    col = "blue",
    linetype = "dashed") +
  geom_vline(
    data = percentile_intervals,
    aes(
      xintercept = .upper),
    col = "blue",
    linetype = "dashed") +
  labs(
    title = "Bootstrap Distributions of Coefficients",
    x = "Coefficient Estimate",
    y = "Count") +
  theme_classic()
```
